Decoding the Impact: Statistical Analysis of Factors Contributing to NFL Injuries

### 2024-04-19 paper.qmd
small changes

---
title: "Decoding the Impact: Statistical Analysis of Factors Contributing to NFL Injuries"
author: "Vincent Xie"
toc: false
number-sections: false
execute:
    echo: false
highlight-style: pygments
format: 
  html: 
    code-fold: true
    html-math-method: katex
  pdf: 
    fontsize: 12pt
    geometry: 
      - top=30mm
      - left=20mm
header-includes:
  - \usepackage{setspace}
  - \doublespacing
##  docx: default
---
```{python}
# Packages
import pandas as pd
import matplotlib.pyplot as plt
import seaborn as sns
```

@@

Prior research in sports medicine and athletic injuries has primarily focused on immediate causes and rehabilitation of injuries. Within the NFL, the annual health and safety reports provide aggregate data on injury types and frequencies, which serve as a baseline for this study. However, there is a lack of comprehensive analysis that integrates environmental conditions, play strategies, and player-specific variables to predict and analyze injury occurrences comprehensively.

@@

```{python}
# Load the datasets
injury = pd.read_csv('data/InjuryRecord.csv')
plays = pd.read_csv('data/PlayList.csv')

# Summarize missing data for InjuryRecord.csv
injury_missing_summary = injury.isnull().sum()
injury_missing_summary = injury_missing_summary[injury_missing_summary > 0]

@@

# Modeling

@@

@@

>>>>

---
title: "Decoding the Impact: Statistical Analysis of Factors Contributing to NFL Injuries"
author: "Vincent Xie"
toc: false
number-sections: false
execute:
    echo: false
highlight-style: pygments
format: 
  html: 
    code-fold: true
    html-math-method: katex
  pdf: 
    fontsize: 12pt
    geometry: 
      - top=30mm
      - left=20mm
header-includes:
  - \usepackage{setspace}
  - \doublespacing
##  docx: default
---
```{python}
# Packages
import pandas as pd
import matplotlib.pyplot as plt
import seaborn as sns
import statsmodels.api as sm
from scipy.stats import chi2_contingency
```

@@

Prior research in sports medicine and athletic injuries has primarily focused on immediate causes and rehabilitation of injuries. Within the NFL, the annual health and safety reports provide aggregate data on injury types and frequencies, which serve as a baseline for this study. However, there is a lack of comprehensive analysis that integrates environmental conditions, play strategies, and player-specific variables to predict and analyze injury occurrences comprehensively. There is also research on how the surface type affects the frequency in lower extremity injuries. The results show a positive correlation between the two factors. This paper will also perform statistical analysis on how specific factors affect the frequency of lower extremity injuries. It may support the claims of prior research or provide new insight on other causes that may contribute to these injuries.

@@

```{python}
# Load the datasets
injury = pd.read_csv('data/InjuryRecord.csv')
plays = pd.read_csv('data/PlayList.csv')
```

```{python}
# Summarize missing data for InjuryRecord.csv
injury_missing_summary = injury.isnull().sum()
injury_missing_summary = injury_missing_summary[injury_missing_summary > 0]

@@

```{python}
# Fill missing 'PlayKey' values with 'GameID' concatenated with '-1'
injury['PlayKey'] = injury.apply(
    lambda row: f"{row['GameID']}-1" if pd.isnull(row['PlayKey']) else row['PlayKey'],
    axis=1
)

# Save the modified dataframe if needed
injury.to_csv('data/InjuryRecord.csv', index=False)
```
# Modeling

@@

```{python}
# Create a set of unique keys from InjuryRecord which indicates injuries
injury_keys = set(injury['PlayKey'])

# Create a binary variable in PlayList that indicates whether an injury occurred
plays['InjuryOccurred'] = plays['PlayKey'].isin(injury_keys).astype(int)

# Create a contingency table of injuries by surface type
contingency_table = pd.crosstab(plays['InjuryOccurred'], injury['Surface'])

# Perform the Chi-Square test
chi2, p, dof, expected = chi2_contingency(contingency_table)

print(f"Chi-Square Statistic: {chi2}, p-value: {p}")
```

@@

# Websites for reference

+ https://pubmed.ncbi.nlm.nih.gov/30452873/
